Commutative character sheaves and geometric types for supercuspidal representations

### 2017-05-04 Geometric types.tex
I'm running out of time so will commit now. But I haven't finished addressing all David's good comments.

We now summarize the sections of the paper in more detail.
%
In Section \ref{sec:defs}, we recall the category $\CS(G)$ from \cite{cunningham-roe:13a} and note that it still makes sense when $G$ is not commutative.  
%
We focus on the case of commutative $G$ in Section \ref{sec:comcom},
giving the definition of a commutative character sheaf and proving our first main theorem, that
$\TrFrob{G} : \CSiso{G} \to G(k)^*$ induces an isomorphism on $\CCSiso{G}$.
%
Passing to the case that $G$ is non-commutative, we give the definition of and main results about commutative character sheaves in Section \ref{sec:noncom}.  
We note that we should only consider character sheaves that arise via pullback from $G\ab$ in order to eliminate those that have nontrivial restriction to the derived subgroup.  
This observation underlies the definition of commutative character sheaves for non-commutative $G$.  
We state our second main result, Theorem~\ref{thm:geo}, that pullback along the abelianization map defines an equivalence of categories $\CCS(G) \to \CCS(G\ab)$.
In Section \ref{ssec:obmor}, we use Galois cohomology to describe the relationship between $G(k)^*$ and $G\ab(k)^*$.  
We also compute the automorphism groups in $\CCS(G)$.
Then in Section \ref{ssec:proofs}, we give proofs of the results in Section \ref{sec:noncom}, which require a development of equivariant linear character sheaves. \todo{Fix this.}
% 
In Section~\ref{sec:types} we use Theorem~\ref{thm:geo} to geometrize types for supercuspidal representations of $p$-adic groups, in a sense made precise in Theorem~\ref{thm:geotypes}. 
% is given in Section~\ref{ssec:geotypes}. 
 %Specifically, we show that if $C$ is a compact $p$-adic group and $\rho$ is a representation of $C$ and $(C,\rho)$ is a type for a tamely ramified $p$-adic group then there is a  group scheme $G$ over $\Fq$ such that $G(\Fq) = C$ and a virtual Weil sheaf $\mathcal{A}$ on $G$ such that $\trFrob{\cs{A}} = \trace(\rho)$.
As preparation for the proof, we review some facts about the Heisenberg-Weil representation and its geometrization, in Section~\ref{ssec:Jacobi}.
Then, in Section~\ref{ssec:review}, we review Yu's theory of types and his study of smooth integral models.  
These elements are pulled together in Section~\ref{ssec:geotypes}, where the proof Theorem~\ref{thm:geotypes} is given.

@@

%
We will define commutative character sheaves as a subcategory of $\CSab(G)$.
%We may now define commutative character sheaves on $G$.
%
Every $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines an isomorphism $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ as follows.
Let $i : G \to G$ be inversion and $c : G\times G\to G\der$ be the commutator map, defined by $c(x,y)= xyx^{-1}y^{-1}$.
Both are smooth morphisms of $\Fq$-schemes.
Set $m' = i \circ m \circ \theta$; we have $j\circ c = m \circ (m \times m')$. \todo{$j$ is not defined, and the needed definition conflicts with the definition of $i$ and $j$ given later.}
Then $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ by the diagram of isomorphisms below.
\begin{equation}
\begin{tikzcd}
\arrow[equal]{d} c^* (\cs{L}\vert_{G\der}) \arrow{r}{c^*(\beta)} 
	&  c^*((\EE)_{G\der}) \arrow[equal]{d} \\
\arrow{d}[swap]{j\circ c = m \circ (m \times m')} c^* j^* \cs{L} 
	&   (\EE)_{G\times G} \\
(m \times m')^* m^* \cs{L} \arrow{d}[swap]{(m \times m')^*(\mu)} 
	&  m^*\cs{L} \otimes \theta^* m^* i^* \cs{L}  \arrow[dashed]{u}[swap]{\gamma'} \\
(m \times m')^* (\cs{L} \boxtimes \cs{L}) \arrow[equal]{r} 
	& m^*\cs{L} \otimes (m')^*\cs{L} \arrow{u}[swap]{m' = i\circ m\circ \theta} 
\end{tikzcd}
\end{equation}
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations; so, for example, the middle left isomorphism comes from $(m\times m')^* m^* \iso c^* j^*$ since $j\circ c = m \circ (m \times m')$.
Using the monoidal structure of the category of Weil local systems on $G\times G$, the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ defines an isomorphism
\[
m^*\cs{L} \to (\theta^* m^* i^*\cs{L})^\vee.
\]
Applying the canonical isomorphisms $(\theta^* m^* i^*\cs{L})^\vee \iso \theta^* m^* i^* (\cs{L}^\vee)$ and $i^*(\cs{L}^\vee) \cong \cs{L}$, this map provides the promised isomorphism
\[
\begin{tikzcd}
\gamma : m^*\cs{L} \arrow{r} & \theta^* m^* \cs{L}.
\end{tikzcd}
\]

\begin{definition}\label{def:CCS}
The category $\CCS(G)$ of commutative character sheaves on $G$ is the full subcategory of $\CSab(G)$ consisting of triples $(\cs{L},\mu,\beta)$ such that the following diagram of Weil sheaves on $G \times G$ commutes:
  \[
  \begin{tikzcd}[row sep=30]
   m^*\cs{L} \arrow{d}[swap]{\gamma} \arrow{r}{\mu} & \cs{L}\boxtimes\cs{L} \arrow{d}{\vartheta}\\
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} &  \theta^*(\cs{L}\boxtimes\cs{L}).
  \end{tikzcd}
  \]
Here $\gamma : m^*\cs{L} \to \theta^* m^* \cs{L}$ is the isomorphism built from $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ above. 
\end{definition}

@@

\begin{lemma}\label{lemma:morinCSab}
If $(\cs{L},\mu,\beta)\in \CS\ab(G)$ then $\mu  : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$, $\gamma : m^*\cs{L} \to  \theta^*(m^*\cs{L})$ and $\vartheta: \cs{L}\boxtimes\cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$ are morphisms in $\CS\ab(G\times G)$.
\end{lemma}
\begin{proof}
Set $G^2 = G\times G$. Define $m^2 : G^2\times G^2 \to G^2$ by $m^2(g_1,g_2,g_1',g_2') = (g_1g_1',g_2g_2')$. Also define $p^2_i : G^2\times G^2 \to G^2$ by $p^2_i(g_1,g_2,g_1,'g_2') = (g_i,g_i')$.
First we show that $m^*\cs{L}$ is an object in $\CS(G^2)$ by equipping it with an isomorphism $\mu_m^2: (m^2)^* (m^*\cs{L}) \to m^*\cs{L} \boxtimes m^*\cs{L}$ defined by the diagram below.
\[
\begin{tikzcd}
(m^2)^* (m^*\cs{L}) \arrow{d}{(m^2)^*\mu} \arrow{rr}{\mu_m^2} && m^*\cs{L} \boxtimes m^*\cs{L}\\
 (m^2)^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & (m^2)^*(p_1)^*\cs{L} \otimes (m^2)^*(p_2)^*\cs{L} \arrow{r} & (p^2_1)^*m^*\cs{L} \otimes (p^2_2)^*m^*\cs{L} \arrow[equal]{u}
\end{tikzcd}
\] 
The pair $(m^*\cs{L},\mu_m^2)$ satisfies the conditions appearing in \cite{cunningham-roe:13a}*{Def. 1.1}.
The restriction of $m^*\cs{L}$ to $G^2\der= G\der\times G\der$ is canonically isomorphic to $(\EE)_{G^2\der}$ by
\[
\begin{tikzcd}
(m^*\cs{L})\vert_{G^2\der} \arrow{d}{\mu\vert_{G^2\der}} \arrow{r}{\beta^2_m} & (\EE)_{G^2\der}\\
(\cs{L}\boxtimes\cs{L})\vert_{G^2\der} \arrow[equal]{r} & (\cs{L}\vert_{G\der})\boxtimes(\cs{L}\vert_{G\der}) \arrow{u}{\beta\boxtimes\beta}.
\end{tikzcd}
\]
This shows that $(m^*\cs{L},\mu_m^2,\beta_m^2) \in \CS\ab(G)$. 
Similar work defines $(\cs{L}\boxtimes\cs{L},\mu_\boxtimes^2,\beta_\boxtimes^2) \in \CS\ab(G)$.
By construction, $\mu: m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ is a morphism in $\CS\ab(G^2)$.
Similar work shows that $\gamma : m^*\cs{L} \to  \theta^*(m^*\cs{L})$ and $\vartheta: \cs{L}\boxtimes\cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$ are also morphisms in $\CS\ab(G^2)$.
\end{proof}

In order to study commutative character sheaves, we need some auxiliary categories.
\begin{itemize}
\item Let $\Loc(G)$ and $\Loc(G\ab)$ be the categories of local systems on $G$ and $G\ab$, respectively.
\item Let $\Loc_{G\der}(G)$ be the category of $G\der$-equivariant local systems on $G$, whose definition we now recall.
Let $m : G\times G\to G$ be multiplication in $G$, $n : G\der\times G\to G$ be its restriction to $G\der\times G$,
$p : G\der\times G \to G$ projection to the second component, and $s: G \to G\der\times G$ be given by $s(g) = (1,g)$.
Then objects in $\Loc_{G\der}(G)$ are pairs $(\cs{L},\nu)$ where $\cs{L}$ is a local system on $G$ and
$\nu : n^*\cs{L} \to p^*\cs{L}$ is an isomorphism such that the following two diagrams commute.
\[
\begin{tikzcd}
s^*n^*\cs{L} \arrow{r}{s^*(\nu)} & s^*p^*\cs{L}\\
\cs{L} \arrow[equal]{u}{n\circ s = \id_G}  \arrow{r}{\id_{\cs{L}}} & \cs{L} \arrow[equal]{u}[swap]{p\circ s = \id_G},
\end{tikzcd}
\]
and with $n_1(h_1,h_2,g) = (h_1h_2,g)$, $n_2(h_1,h_2,g) = (h_1,n(h_2,g))$ and $p_2(h_1,h_2,g) = (h_2,g)$,
\[
\begin{tikzcd}
& \arrow[equal]{dl}{n\circ n_2=n\circ n_1} n_2^*n^*\cs{L} \arrow{rr}{n_2^*(\nu)} && n_2^*p^*\cs{L} \arrow[equal]{dr}[swap]{p \circ n_2 = n\circ p_2}& \\
n_1^*n^*\cs{L} \arrow{dr}[swap]{n_1^*(\nu)} &&&& \arrow{dl}{p_2^*(\nu)} p_2^*n^*\cs{L}\\
& n_1^*p^*\cs{L} \arrow[equal]{rr}{p\circ n_1 = p\circ p_2} && p_2^*p^*\cs{L} .
\end{tikzcd}
\]
Morphisms are morphisms in $\Loc(G)$ commuting with $\nu$.
\item Let $\CSab'(G)$ be the category of triples $(\cs{L},\nu,\mu)$ where $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\nu)\in \Loc_{G\der}(G)$; thus, $\mu : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ and $\nu :  n^*\cs{L} \to p^*\cs{L}$ are isomorphisms.
Morphisms are morphisms in $\Loc(G)$ commuting with $\mu$ and $\nu$. 
\end{itemize}
We relate $\CSab(G)$ to $\CS(G\ab)$ via the quotient map $q : G \to G\ab$.

@@

Let $i : G\der \to G$ be the kernel of $q:G \to G\ab$ and define $j : G\der \to G\times G$ by $j(h) = (i(h),1_G)$. 
For $(\cs{L},\mu,\nu)\in \CSab'(G)$, $j^*(\nu) : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism. \todo{For $j^*(\nu)$ to make sense, need $j : G\der \to G\der \times G$}
This defines a functor
\[
\begin{array}{rcl}
\CSab'(G) &\to& \CSab(G) \\
(\cs{L},\mu,\nu) &\mapsto& (\cs{L},\mu,j^*(\nu)).
\end{array}
\]
Define $k: G\der\times G\to G\times G$ by $k(h,g) = (i(h),g)$.
Then for $(\cs{L},\mu,\beta) \in \CSab(G)$, define $\nu : n^*\cs{L} \to p_n^*\cs{L}$ \todo{I assume this should be $p^*\cs{L}$?} by
\[
\begin{tikzcd}
& n^* \cs{L} \arrow{dl}{m\circ k = n} \arrow{r} & p_n^*\cs{L} &  \\ 
k^* m^* \cs{L} \arrow{dr}{k^*(\mu)} &&& (\EE)_{G\der} \boxtimes \cs{L} \arrow[equal]{ul} \\
& k^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & \cs{L}\vert_{G\der}\boxtimes \cs{L} \arrow{ur}{\beta\times 1}
\end{tikzcd}
\]
Then $(\cs{L},\mu,\nu)\in \CSab'(G)$ maps to $(\cs{L},\mu,\beta)$.
It now follows that $\CSab'(G) \to \CSab(G)$ is an equivalence. \todo{Is it easy to see why the morphisms that commute with $\mu$ and $\beta$ are the same as those commuting with $\mu$ and $\nu$?  Is this straight from the definition of the functor?}
\end{proof}

@@

\begin{proof}
Notation as in the proof of Proposition~\ref{prop:CSGabab}.
Consider the following diagram.
%\[
%\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
%1 \arrow{r} & H \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
%\end{tikzcd}
%\]
%gives
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{q\ab^*} \\
& \CSab(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]

@@

Conversely, suppose $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\mu) \iso q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$.
Then 
\[
i^*(\cs{L},\mu) \iso i^*q^*(\cs{L}\ab,\mu\ab)
\]
 in $\CS(G\der)$.
Since $q\circ i = 1_{G\ab}$, it follows that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$ in $\CS(G\der)$.
\end{proof}

>>>>

We now summarize the sections of the paper in more detail.
%
In Section \ref{sec:defs}, we recall the category $\CS(G)$ from \cite{cunningham-roe:13a} and note that it still makes sense when $G$ is not commutative.  
%
We focus on the case of commutative $G$ in Section \ref{sec:comcom},
giving the definition of a commutative character sheaf and proving our first main theorem, that
$\TrFrob{G} : \CSiso{G} \to G(k)^*$ induces an isomorphism on $\CCSiso{G}$.
%
Passing to the case that $G$ is non-commutative, we give the definition of and main results about commutative character sheaves in Section \ref{sec:noncom}.  
We note that we should only consider character sheaves that arise via pullback from $G\ab$ in order to eliminate those that have nontrivial restriction to the derived subgroup.  
This observation underlies the definition of commutative character sheaves for non-commutative $G$.  
We state our second main result, Theorem~\ref{thm:geo}, that pullback along the abelianization map defines an equivalence of categories $\CCS(G) \to \CCS(G\ab)$.
In Section \ref{ssec:obmor}, we use Galois cohomology to describe the relationship between $G(k)^*$ and $G\ab(k)^*$.  
We also compute the automorphism groups in $\CCS(G)$.
%Then in Section \ref{ssec:proofs}, we give proofs of the results in Section \ref{sec:noncom}, which require a development of equivariant linear character sheaves. \todo{Fix this.}
% 
In Section~\ref{sec:types} we use Theorem~\ref{thm:geo} to geometrize types for supercuspidal representations of $p$-adic groups, in a sense made precise in Theorem~\ref{thm:geotypes}. 
% is given in Section~\ref{ssec:geotypes}. 
 %Specifically, we show that if $C$ is a compact $p$-adic group and $\rho$ is a representation of $C$ and $(C,\rho)$ is a type for a tamely ramified $p$-adic group then there is a  group scheme $G$ over $\Fq$ such that $G(\Fq) = C$ and a virtual Weil sheaf $\mathcal{A}$ on $G$ such that $\trFrob{\cs{A}} = \trace(\rho)$.
As preparation for the proof, we review some facts about the Heisenberg-Weil representation and its geometrization, in Section~\ref{ssec:Jacobi}.
Then, in Section~\ref{ssec:review}, we review Yu's theory of types and his study of smooth integral models.  
These elements are pulled together in Section~\ref{ssec:geotypes}, where the proof Theorem~\ref{thm:geotypes} is given.

@@

%
We will define commutative character sheaves as a subcategory of $\CSab(G)$.
%We may now define commutative character sheaves on $G$.
%
Every $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines an isomorphism $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ as follows.
Let $i : G \to G$ be inversion and $c : G\times G\to G\der$ be the commutator map, defined by $c(x,y)= xyx^{-1}y^{-1}$.
Both are smooth morphisms of $\Fq$-schemes.
Set $m' = i \circ m \circ \theta$ and let $j\der : G\der\to G$ be inclusion; then $j\der\circ c = m \circ (m \times m')$.
% \todo{$j$ is not defined, and the needed definition conflicts with the definition of $i$ and $j$ given later.}
Then $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ by the diagram of isomorphisms below.
\begin{equation}
\begin{tikzcd}
\arrow[equal]{d} c^* (\cs{L}\vert_{G\der}) \arrow{r}{c^*(\beta)} 
	&  c^*((\EE)_{G\der}) \arrow[equal]{d} \\
\arrow{d}[swap]{j\der\circ c = m \circ (m \times m')} c^* j\der^* \cs{L} 
	&   (\EE)_{G\times G} \\
(m \times m')^* m^* \cs{L} \arrow{d}[swap]{(m \times m')^*(\mu)} 
	&  m^*\cs{L} \otimes \theta^* m^* i^* \cs{L}  \arrow[dashed]{u}[swap]{\gamma'} \\
(m \times m')^* (\cs{L} \boxtimes \cs{L}) \arrow[equal]{r} 
	& m^*\cs{L} \otimes (m')^*\cs{L} \arrow{u}[swap]{m' = i\circ m\circ \theta} 
\end{tikzcd}
\end{equation}
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations; so, for example, the middle left isomorphism comes from $(m\times m')^* m^* \iso c^* j\der^*$ since $j\der\circ c = m \circ (m \times m')$.
Using the monoidal structure of the category of Weil local systems on $G\times G$, the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ defines an isomorphism
\[
m^*\cs{L} \to (\theta^* m^* i^*\cs{L})^\vee.
\]
Applying the canonical isomorphisms $(\theta^* m^* i^*\cs{L})^\vee \iso \theta^* m^* i^* (\cs{L}^\vee)$ and $i^*(\cs{L}^\vee) \cong \cs{L}$, this map provides the promised isomorphism
\[
\begin{tikzcd}
\gamma : m^*\cs{L} \arrow{r} & \theta^* m^* \cs{L}.
\end{tikzcd}
\]

\begin{definition}\label{def:CCS}
The category $\CCS(G)$ of commutative character sheaves on $G$ is the full subcategory of $\CSab(G)$ consisting of triples $(\cs{L},\mu,\beta)$ such that the following diagram of Weil sheaves on $G \times G$ commutes:
  \[
  \begin{tikzcd}[row sep=30]
   m^*\cs{L} \arrow{d}[swap]{\gamma} \arrow{r}{\mu} & \cs{L}\boxtimes\cs{L} \arrow{d}{\vartheta}\\
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} &  \theta^*(\cs{L}\boxtimes\cs{L}).
  \end{tikzcd}
  \]
Here $\gamma : m^*\cs{L} \to \theta^* m^* \cs{L}$ is the isomorphism built from $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ as above. 
\end{definition}

@@

\begin{lemma}\label{lemma:morinCSab}
If $(\cs{L},\mu,\beta)\in \CS\ab(G)$ then $\mu  : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$, $\gamma : m^*\cs{L} \to  \theta^*(m^*\cs{L})$ and $\vartheta: \cs{L}\boxtimes\cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$ are morphisms in $\CS\ab(G\times G)$.
\end{lemma}
\begin{proof}
Set $G^2 = G\times G$. Define $m^2 : G^2\times G^2 \to G^2$ by $m^2(g_1,g_2,g_1',g_2') = (g_1g_1',g_2g_2')$. Also define $p^2_i : G^2\times G^2 \to G^2$ by $p^2_i(g_1,g_2,g_1,'g_2') = (g_i,g_i')$.
First we show that $m^*\cs{L}$ is an object in $\CS(G^2)$ by equipping it with an isomorphism $\mu_m^2: (m^2)^* (m^*\cs{L}) \to m^*\cs{L} \boxtimes m^*\cs{L}$ defined by the diagram below.
\[
\begin{tikzcd}
(m^2)^* (m^*\cs{L}) \arrow{d}{(m^2)^*\mu} \arrow{rr}{\mu_m^2} && m^*\cs{L} \boxtimes m^*\cs{L}\\
 (m^2)^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & (m^2)^*(p_1)^*\cs{L} \otimes (m^2)^*(p_2)^*\cs{L} \arrow{r} & (p^2_1)^*m^*\cs{L} \otimes (p^2_2)^*m^*\cs{L} \arrow[equal]{u}
\end{tikzcd}
\] 
The pair $(m^*\cs{L},\mu_m^2)$ satisfies the conditions appearing in \cite{cunningham-roe:13a}*{Def. 1.1}.
The restriction of $m^*\cs{L}$ to $G^2\der= G\der\times G\der$ is canonically isomorphic to $(\EE)_{G^2\der}$ by
\[
\begin{tikzcd}
(m^*\cs{L})\vert_{G^2\der} \arrow{d}{\mu\vert_{G^2\der}} \arrow{r}{\beta^2_m} & (\EE)_{G^2\der}\\
(\cs{L}\boxtimes\cs{L})\vert_{G^2\der} \arrow[equal]{r} & (\cs{L}\vert_{G\der})\boxtimes(\cs{L}\vert_{G\der}) \arrow{u}{\beta\boxtimes\beta}.
\end{tikzcd}
\]
This shows that $(m^*\cs{L},\mu_m^2,\beta_m^2) \in \CS\ab(G^2)$. 
Similar work defines $(\cs{L}\boxtimes\cs{L},\mu_\boxtimes^2,\beta_\boxtimes^2) \in \CS\ab(G^2)$.
By construction, $\mu: m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ is a morphism in $\CS\ab(G^2)$.
Similar work shows that $\gamma : m^*\cs{L} \to  \theta^*(m^*\cs{L})$ and $\vartheta: \cs{L}\boxtimes\cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$ are also morphisms in $\CS\ab(G^2)$.
\end{proof}

In order to study commutative character sheaves, we need some auxiliary categories.
\begin{itemize}
\item Let $\Loc(G)$ and $\Loc(G\ab)$ be the categories of local systems on $G$ and $G\ab$, respectively.
\item Let $\Loc_{G\der}(G)$ be the category of $G\der$-equivariant local systems on $G$, whose definition we now recall.
Let $n : G\der\times G\to G$ be the restriction of $m : G\times G\to G$ to $G\der\times G$,
let $p_n : G\der\times G \to G$ be projection to the second component, and let $s: G \to G\der\times G$ be given by $s(g) = (1,g)$.
Objects in $\Loc_{G\der}(G)$ are pairs $(\cs{L},\nu)$ where $\cs{L}$ is a local system on $G$ and
$\nu : n^*\cs{L} \to p^*\cs{L}$ is an isomorphism such that the following two diagrams commute.
\[
\begin{tikzcd}
s^*n^*\cs{L} \arrow{r}{s^*(\nu)} & s^*p^*\cs{L}\\
\cs{L} \arrow[equal]{u}{n\circ s = \id_G}  \arrow{r}{\id_{\cs{L}}} & \cs{L} \arrow[equal]{u}[swap]{p\circ s = \id_G},
\end{tikzcd}
\]
and with $n_1(h_1,h_2,g) = (h_1h_2,g)$, $n_2(h_1,h_2,g) = (h_1,n(h_2,g))$ and $p_2(h_1,h_2,g) = (h_2,g)$,
\[
\begin{tikzcd}
& \arrow[equal]{dl}{n\circ n_2=n\circ n_1} n_2^*n^*\cs{L} \arrow{rr}{n_2^*(\nu)} && n_2^*p^*\cs{L} \arrow[equal]{dr}[swap]{p \circ n_2 = n\circ p_2}& \\
n_1^*n^*\cs{L} \arrow{dr}[swap]{n_1^*(\nu)} &&&& \arrow{dl}{p_2^*(\nu)} p_2^*n^*\cs{L}\\
& n_1^*p^*\cs{L} \arrow[equal]{rr}{p\circ n_1 = p\circ p_2} && p_2^*p^*\cs{L} .
\end{tikzcd}
\]
Morphisms are morphisms in $\Loc(G)$ commuting with $\nu$.
\item Let $\CSab'(G)$ be the category of triples $(\cs{L},\nu,\mu)$ where $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\nu)\in \Loc_{G\der}(G)$; thus, $\mu : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ and $\nu :  n^*\cs{L} \to p^*\cs{L}$ are isomorphisms.
Morphisms in $\Loc_{G\der}(G)$ are morphisms in $\Loc(G)$ commuting with $\mu$ and $\nu$. 
\end{itemize}

\begin{lemma}\label{lemma:descent}
\end{lemma}

We may now relate $\CSab(G)$ to $\CS(G\ab)$ via the quotient map $q : G \to G\ab$.

@@

Let $i\der : G\der \to G$ be the kernel of $q:G \to G\ab$ and define $j : G\der \to G\der\times G$ by $j(h) = (i\der(h),1_G)$. 
For $(\cs{L},\mu,\nu)\in \CSab'(G)$, $j^*(\nu) : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism. \todo{For $j^*(\nu)$ to make sense, need $j : G\der \to G\der \times G$}
This defines a functor
\[
\begin{array}{rcl}
\CSab'(G) &\to& \CSab(G) \\
(\cs{L},\mu,\nu) &\mapsto& (\cs{L},\mu,j^*(\nu)).
\end{array}
\]
Define $k: G\der\times G\to G\times G$ by $k(h,g) = (i\der(h),g)$.
Then for $(\cs{L},\mu,\beta) \in \CSab(G)$, define $\nu : n^*\cs{L} \to p_n^*\cs{L}$ \todo{I assume this should be $p^*\cs{L}$?} by
\[
\begin{tikzcd}
& n^* \cs{L} \arrow{dl}{m\circ k = n} \arrow{r} & p_n^*\cs{L} &  \\ 
k^* m^* \cs{L} \arrow{dr}{k^*(\mu)} &&& (\EE)_{G\der} \boxtimes \cs{L} \arrow[equal]{ul} \\
& k^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & \cs{L}\vert_{G\der}\boxtimes \cs{L} \arrow{ur}{\beta\times 1}
\end{tikzcd}
\]
Then $(\cs{L},\mu,\nu)\in \CSab'(G)$ maps to $(\cs{L},\mu,\beta)$.
It now follows that $\CSab'(G) \to \CSab(G)$ is an equivalence. \todo{Is it easy to see why the morphisms that commute with $\mu$ and $\beta$ are the same as those commuting with $\mu$ and $\nu$?  Is this straight from the definition of the functor?}
\end{proof}

@@

\begin{proof}
Notation as in the proof of Proposition~\ref{prop:CSGabab}.
Consider the following diagram.
%\[
%\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
%1 \arrow{r} & H \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
%\end{tikzcd}
%\]
%gives
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i\der^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{q\ab^*} \\
& \CSab(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]

@@

Conversely, suppose $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\mu) \iso q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$.
Then 
\[
i\der^*(\cs{L},\mu) \iso i\der^*q^*(\cs{L}\ab,\mu\ab)
\]
 in $\CS(G\der)$.
Since $q\circ i\der = 1$, it follows that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$ in $\CS(G\der)$.
\end{proof}
